\section{Voice Mode and spoken conversational practice}

\subsection{An exploratory German case study}

A small exploratory case study involved two co-authors of this report, Sarah
Wright and Manny Rayner. Sarah was preparing for an extended trip to Europe,
including many weeks in German-speaking countries. She had studied German at
school and retained substantial passive knowledge, but lacked confidence in
using German spontaneously in practical situations. She asked Manny whether he
could act as a conversation partner. Manny declined, judging that his own
spoken German was not strong enough for that role, and instead suggested trying
ChatGPT Pro in Voice Mode.

The first session was sufficiently successful that Sarah chose to continue
with approximately weekly German sessions for about eight weeks, normally
lasting 60--90 minutes, despite substantial commitments elsewhere. The early
sessions focused on practical travel situations, including checking into a
campsite, visiting a pharmacy, buying food, asking for help and dealing with
transport. Later sessions became increasingly open-ended and personally
meaningful. This was not a controlled study and does not demonstrate measured
learning gains, but Sarah's sustained voluntary use provides practical evidence
that she found the interaction useful and motivating.

The sessions were deliberately simple. Sarah and Manny sat together and used
German as the main language of interaction with the publicly available
ChatGPT Pro Voice Mode system. No special lesson script or experimental
interface was used. The evidence is anecdotal: participant recollections,
visible conversation history where available, and post-session summaries.
Branislav Bédi attended one early session and followed the later work through
these summaries. The case should therefore be read as an ecologically realistic
exploration of a readily reproducible practice, not as a controlled evaluation.

\subsection{Observed practical value and current limitations}

Voice Mode functioned as a patient, flexible conversation partner rather than
a specialised CALL tutor. Its main apparent value was confidence-building:
Sarah could continue speaking despite uncertainty, ask for vocabulary as
needed and move from rehearsed travel scenarios towards freer interaction.
When she lacked a word or phrase, the system could usually supply it within the
ongoing conversation rather than interrupting the interaction with a formal
exercise. For example, practical vocabulary concerning a bracelet scam near
the Eiffel Tower and vocabulary about cat food and cat treats arose because
Sarah wanted to describe real events and interests. The assistance was thus
situated: vocabulary was introduced in order to sustain a meaningful
conversation rather than as a pre-planned list.

The limitations were equally clear. Voice Mode did not provide configurable
correction, reliable speaker attribution, systematic learner tracking or
automatic factual checking. Speech recognition and turn-taking could also be
imperfect, particularly in a two-speaker session involving hesitant or
code-switched learner speech. Most importantly, the participants' experience
was that the quality required for extended, flexible spoken practice depended
on ChatGPT Pro. Attempts to use the lower-cost Plus tier in early 2026 produced
performance too weak for practical use in this setting. This may help explain
why Voice Mode has not generally been recognised as a language-practice tool:
Pro is commonly associated with coding, research and business applications,
and its cost is a serious barrier to widespread learner access.

These observations are consistent with a wider interest in AI chatbots for
language learning \citep{KocSavas2025}, but the present case makes only a
modest claim. A general-purpose spoken AI system can already provide a
low-pressure supplementary space for sustained conversation practice. It is
not yet an affordable, specialised or systematically evaluated CALL tutor.

\subsection{Status and future direction}

Voice Mode is not currently an implemented C-LARA-2 feature, and the account
in this section is an initial anecdotal exploration rather than evidence of
measured learning gains. It nevertheless identifies a promising direction:
spoken interaction can be personally meaningful, learner-driven and capable of
supplying immediate conversational scaffolding, while the current general-
purpose system lacks the pedagogical organisation needed for a fuller CALL
offering.

Section~11 returns to the future question of how free spoken interaction might
be connected to C-LARA-2's multimodal texts, picture resources,
learner-relevant vocabulary and post-session processing. The aim would not be
to replace teachers or ordinary human interaction, but to connect additional
spoken practice with inspectable learning materials and human pedagogical
judgement.
