\section{Future directions}

C-LARA-2 is not a finished system but a rapidly developing platform. The
priorities outlined here are intended to guide work over the next six to eight
months. They combine problems revealed by the current implementation with
requests from the communities, teachers and learners for whom the platform is
being developed. We expect the detailed order to change as experience and
feedback accumulate, but the following directions are already sufficiently
clear to inform planning.

\subsection{Improving linguistic annotation through curated examples}

The most urgent priority concerns linguistic annotation. C-LARA-2 has a more
modular and inspectable architecture than its predecessor C-LARA, but its
performance on central annotation tasks---segmentation, lemmatisation,
glossing, multi-word expressions and related operations---has so far sometimes
appeared weaker. Our working explanation is not that the new architecture is
incapable of good results, but that the prompts used by the new pipeline are
not yet as effective as those developed for C-LARA. The older prompts were
hand-written and refined over time; the C-LARA-2 prompts have, consistently
with the project's general development model, been AI-generated.

A simple response would be to import the older prompts or to rewrite them by
hand. This would probably improve the immediate output, but it would evade the
more important question: how can an AI-assisted project systematically acquire
and improve high-quality prompts? Our current priority is therefore to develop
a repeatable example-curation method in which linguists supply moderately small
sets of carefully chosen examples, including difficult and representative
cases, and the AI derives and evaluates prompt variants from them. The work
will use deterministic checks to ensure that proposed annotations preserve the
input and obey the required schema, alongside AI review and targeted human
audit. Held-out examples and versioned experiment records should make it
possible to test whether a change genuinely improves performance rather than
merely producing a more persuasive-looking prompt. This is both a practical
need and an important test of the wider claim that AI can maintain the
project's technical knowledge under human supervision.

\subsection{More community-controlled image workflows}

The quality of generated images is a less acute but highly visible concern.
It is particularly important in the Indigenous-language contexts central to the
project, where an image can be culturally inappropriate, depict an item in an
unfamiliar way, or simply fail to represent what a community contributor had
in mind. The existing review and regeneration workflow already provides useful
controls (Section~5), but feedback suggests two complementary extensions.

First, contributors should be able to upload a reference image and ask the
system to adapt it to the selected project style. This would preserve a
locally meaningful visual starting point while retaining stylistic coherence
across a text or picture dictionary. Second, the image flow should include an
AI-based analysis and quality-control pass. The analysis would compare a
candidate image with its prompt, relevant text and, where appropriate,
community guidance. It could then support a transparent decision to retain the
image, choose among alternatives, regenerate it, or revise the prompt before
trying again. Neither mechanism removes the need for human review: their
purpose is to make repeated image generation more targeted and less laborious,
while giving community contributors stronger control over what is accepted.

\subsection{Human-recorded audio and rapid alignment}

Human-recorded audio is another major requirement, especially for languages
for which available text-to-speech output is absent or unsuitable. One
near-term route is a simple recording interface that presents the segments or
lexical items required by a project and lets a contributor record or rerecord
them in the browser. Recordings can then be reviewed, approved and associated
with the corresponding items in the compiled document. This would provide a
practical community-recording path without requiring contributors to master
audio-processing tools.

A more ambitious route begins with a recording of an entire text. The system
would split the recording at silences, use speech recognition to align the
resulting audio pieces with the text, and construct the corresponding
C-LARA-2 document. A predecessor of this workflow was implemented in LARA and
reported by \citet{RaynerChieraChua2022}: public-domain text, human audio and
translations were combined using silence detection, speech recognition and
alignment, with encouraging results even for substantial literary texts. In a
C-LARA-2 implementation, this could offer a rapid way to turn existing
high-quality recordings into learner resources, while retaining a review step
for imperfect alignments and respecting copyright, consent and community data
sovereignty.

\subsection{Mobile access and spoken interaction}

Many prospective users have asked for at least the learner-facing parts of the
platform to work well on phones and tablets. The first target is therefore a
responsive web experience rather than a separate native application: reading
compiled texts, playing audio, viewing images, using touch-friendly
annotations and completing exercises should all work reliably on ordinary
mobile devices. A later phase can offer a compact, default-driven creation
flow and selected review actions. Dense linguistic annotation and bulk editing
will remain desktop-first until they have been explicitly redesigned for small
screens.

Section~10 described preliminary evidence that high-quality spoken AI
interaction can give learners valuable low-pressure conversational practice.
The corresponding longer-term design opportunity is to integrate a spoken
conversation mode with C-LARA-2 materials. A crude demonstration is already
possible: a user can open C-LARA-2 in one browser window, upload a screenshot
of a page to a ChatGPT Pro Voice Mode conversation in another, and discuss the
page orally. This confirms that texts, images and exercises can provide useful
shared material for spoken interaction, but the manual screenshot-and-switch
procedure is far too cumbersome for ordinary use. Future work should explore
more direct ways to ground spoken conversation in a C-LARA-2 page, and to turn
post-session vocabulary, corrections and summaries into follow-up learning
material.

\subsection{Demand-driven development}

Finally, these larger directions do not exhaust the work required. A living
platform will continue to generate smaller usability problems, language-
specific needs and requests from particular user groups. The roadmap, issue
and how-to mechanisms described in Section~7 are intended to make this work
visible and cumulative: user suggestions can be turned into concrete plans,
implemented changes can be tested and documented, and priorities can be
revised without losing the history of earlier decisions. We therefore expect
the next phase to combine the strategic work above with continuing,
demand-driven refinement of the platform.
