\section{Discussion and conclusion}

This report has documented the first nine months of C-LARA-2 as both a
multimodal CALL platform and an experiment in AI-mediated project development.
We described a modular annotation pipeline and its manual alternatives; an
image-generation workflow built around reviewable intermediate artefacts;
picture dictionaries redesigned around the needs of their community authors;
and a repository that functions as an external project memory through roadmaps,
how-to material, issue records and the project-understanding Assistant. We
also described the deployment work needed to make the platform available,
preliminary evidence about Voice Mode as a spoken-practice resource, and the
near-term programme for annotation quality, images, recorded audio, mobile
access and spoken interaction.

The common theme is not that the AI has replaced the expertise of linguists,
teachers, community contributors or project leads. It has been that an AI
coding assistant can turn human purposes, examples and judgements into
working software, documentation, tests and revised plans at a speed that
changes what a small research team can attempt. The picture-dictionary case is
especially clear: the community collaborators identified that an inherited
interaction model was wrong, while Codex made it practicable to replace the
user-facing layer without abandoning the useful underlying representation.
Likewise, the project's roadmaps, issue files and how-to material are not
passive documentation. They form a maintained memory that supports both
human work and the Assistant's inspectable answers about current behaviour,
known limitations and planned development.

\subsection{What is distinctive about this workflow?}

From one perspective, the project has done nothing mysterious: it has applied
an AI coding assistant consistently to the ordinary tasks involved in an
academic software project. Other groups may well be doing comparable work, and
this report cannot establish how unusual the present workflow is. Two
observations nevertheless suggest that sustained end-to-end use remains less
common than it might soon become.

First, the work would not have been feasible for this team without a ChatGPT
Pro subscription or an equivalent high-end service. The subscription was paid
for personally, rather than supplied through an academic institution. That is
a mundane but consequential constraint: universities and research groups may
be reluctant or unable to fund premium AI access, even when it is the resource
that enables a small team to build, document and operate a substantial system.
The resulting unequal access deserves attention in discussions of AI-enabled
research capacity.

Second, using Codex to draft this report exposed a practical limitation in the
current publication workflow. The environment can inspect images supplied in
conversation, but its pull-request path did not reliably accept binary image
assets. We therefore retained textual figure placeholders in the repository
and prepared the real screenshots separately for Overleaf. This workaround
does not affect the substantive platform claims, but it shows that supposedly
end-to-end AI-assisted repository maintenance still meets tool assumptions
that are poorly matched to the production of an illustrated academic report.
The limitation is also suggestive: if an apparently natural use case is not
well supported, it may be because consistent use of an AI assistant across
software, documentation and publication remains uncommon.

\subsection{From supervised autonomy to progressive delegation}

The organising phrase for the project has been \emph{AI autonomy under human
supervision}. It remains useful, but the deployment work in Section~8 reveals
an important qualification. The human participants formally controlled the
AWS deployment: they selected the goal, supplied credentials where necessary,
ran the instructed commands and could stop the process. Yet in difficult parts
of the work, particularly the read-only Assistant sandbox, they did not fully
understand the relevant techniques or sometimes know that such techniques
existed. In practice they had to decide whether to trust a detailed AI-proposed
solution.

This is not an exceptional logical possibility but a likely pattern of
human--AI work as technical systems become more capable. Whenever an AI can
demonstrably perform a task better than the available human participants,
practical authority may be handed over before human understanding catches up.
The shift need not be abrupt or all-or-nothing. It can take the form of
progressively larger delegated tasks, clear success criteria, bounded choices
among pre-authorised options, logging and retrospective audit, and an explicit
human right to override or halt a process. These mechanisms do not remove the
need to assign human accountability; they make it possible to distinguish
formal authority from the actual distribution of knowledge and operational
control.

The same question extends beyond technical execution to strategy and ethics.
Codex already expands brief user suggestions into detailed plans, identifies
dependencies across several documents and proposes priorities. Modern AI
systems can also offer reasoned ethical advice---for example, recommending an
action, warning against one, and explaining the considerations involved.
People may reasonably be reluctant to delegate fully autonomous ethical
decisions, especially where cultural authority, safety, rights or community
data sovereignty are at stake. But the boundary is not a simple fixed line
between ``ethical advice'' and ``ethical responsibility''. It is an empirical
and governance question: which forms of advice, initiative and bounded choice
are useful; what evidence, escalation rules and review are required; and when
should a human or community authority make the final decision? We regard
AI-maintained strategy plans, ethical-risk advice and explicit escalation
charters as promising subjects for future investigation, rather than as a
reason to assume that strategic work must remain wholly outside AI autonomy.

\subsection{Implications for CALL}

The report does not claim to have demonstrated learning gains, nor does it
claim that C-LARA-2 is ready to provide a complete individual language tutor.
Important work remains on annotation quality, culturally appropriate images,
human audio, mobile access, spoken interaction, cost and classroom-oriented
evaluation. Nevertheless, the combination now becoming available is striking:
AI systems can help create learner-specific multimodal materials, support
community review and authoring, provide extended spoken practice, and maintain
much of the software and project memory required to improve those facilities.

This makes it plausible that CALL is approaching a major advance within the
next two years. The aspiration is not new. The classic ``two sigma'' result
showed the potential of one-to-one tutoring to outperform conventional group
instruction substantially \citep{Bloom1984}; \citet{Mollick2024} has recently
argued that widely available AI systems make forms of scalable, individualised
assistance newly practical. We do not infer a two-sigma effect for language
learning from the present work. Rather, we see a credible path towards
individual language tutors that combine plentiful conversational practice with
materials, feedback and review adapted to an individual learner and to the
knowledge of teachers and communities.

C-LARA-2 is best understood as an early, inspectable contribution to that
path. Its value lies as much in the development method as in its current
features: human collaborators define purposes, standards and boundaries; AI
systems help formulate, implement, test, document and sometimes operationalise
the work; and the repository preserves the evidence from which both groups can
learn. The next phase will test how far this model can be extended while
retaining meaningful community control, pedagogical quality and accountable
human governance.
